---
title: "Collaboration filtering"
author: "Wenxuan Zhu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(skimr)

ratings <- read.csv("/Users/zhuwenxuan/Desktop/365 final project/R_film_data.csv")
```

```{r}
themeEF <- function(){
dark_text = "steelblue4"
light_text = "slategrey"
  theme_minimal()+
  theme(text = element_text(color = light_text, family = "mono"), 
        plot.title = element_text(color = dark_text, family = "mono"), 
        axis.text = element_text(color = light_text, family = "mono"))
}

# scale_fill_manual(values = c("#A4ADAF", "#D4451F", "#01426A"))
```

# Data cleaning / matrix
```{r}
# data cleaning / matrix
# movies <- read.csv("/Users/zhuwenxuan/Desktop/365 final project/ml-latest-small/movies.csv")
# ratings <- read.csv("/Users/zhuwenxuan/Desktop/365 final project/ml-latest-small/ratings.csv")


# dense_movies <- ratings %>% select(movieId) %>% distinct() %>% rowid_to_column()
# ratings <- ratings %>% inner_join(dense_movies) %>% rename(movieIdDense = rowid)
# ratings <- ratings %>% inner_join(movies) %>% select(userId, movieIdDense, rating, title, genres)
# 
# write.csv(ratings, "R_film_data.csv", row.names = FALSE)
```

```{r}
max(ratings$movieIdDense)
max(ratings$userId)
max(ratings$rating)
```

# Data visualization 
```{r Plot: aveScore}
ratings %>% 
  group_by(userId) %>% 
  mutate(aveScore = mean(rating)) %>% 
  ggplot(aes(x = userId, y = aveScore)) +
  geom_point(color = "#01426A") + 
  geom_smooth(method = lm, color = "#D4451F") + 
  themeEF() +
  labs(title = "Average rating score for each user")
```
```{r}
# scale_fill_manual(values = c("#A4ADAF", "#D4451F", "#01426A"))
ratings %>% 
  mutate(year = str_extract(title, "(?<=\\()\\d{4}(?=\\))")) %>% 
  filter(!is.na(year)) %>% 
  group_by(year) %>% 
  summarise(film_count = n()) %>% 
  mutate(mean_year_count = mean(film_count)) %>%  
  ggplot(aes(x = year, y = film_count)) +
  geom_col(col = "#01426A", fill = "#A4ADAF") +
  geom_vline(xintercept = 1985-1902, color = "#D4451F", linetype = "dashed") +  
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  themeEF() +
  scale_x_discrete(breaks = seq(1902, 2018, by = 3)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Number of flims each year",
       subtitle = "Red dash: 1985")
```

# SVT / matrix completion
```{r}
library(filling)
```

# replace NA values with 0
# UImatrix[is.na(UImatrix)] <- 0

# convert UImatrix to a matrix
# UImatrix <- as.matrix(UImatrix)

## subset!
subset_rows <- sample(1:nrow(UImatrix), 100)
UImatrix_sub <- UImatrix[subset_rows, ]

# check for empty columns
UImatrix_sub <- UImatrix_sub[, colSums(is.na(UImatrix_sub)) != nrow(UImatrix_sub)]

subset_cols <- sample(1:ncol(UImatrix_sub), 2000)

# subset the UImatrix
UImatrix_sub <- UImatrix_sub[, subset_cols]

# convert UImatrix to a matrix
UImatrix_sub <- as.matrix(UImatrix_sub)

# find the maximum value ignoring missing values
# max(UImatrix_sub, na.rm = TRUE)


```{r}
UImatrix <- pivot_wider(ratings, id_cols = userId, names_from = movieIdDense, values_from = rating) 
# convert UImatrix to a matrix
# UImatrix <- as.matrix(UImatrix)

## subset!
subset_rows <- sample(1:nrow(UImatrix), 100)
UImatrix_sub <- UImatrix[subset_rows, ]

# check for empty columns
UImatrix_sub <- UImatrix_sub[, colSums(is.na(UImatrix_sub)) != nrow(UImatrix_sub)]
subset_cols <- sample(1:ncol(UImatrix_sub), 2000)

# subset the UImatrix
UImatrix_sub <- UImatrix_sub[, subset_cols]

# get corresponding movie titles
movie_titles <- ratings$title[match(colnames(UImatrix_sub), ratings$movieIdDense)]

# set column names of UImatrix_sub to corresponding movie titles
colnames(UImatrix_sub) <- movie_titles

# convert UImatrix to a matrix
UImatrix_sub <- as.matrix(UImatrix_sub)
```


```{r grid search, cache=TRUE}
# create a sequence of lambda values to try
lambda_seq <- seq(0.001, 1.001, by = 0.05)

# initialize an empty data frame to store the results
results <- data.frame(lambda = numeric(),
                      rmse = numeric())

# loop through each lambda value and compute RMSE
for (lambda in lambda_seq) {
  # apply SVT with current lambda value
  svt_result <- fill.SVT(UImatrix_sub, lambda = lambda, maxiter = 50, tol = 0.01)
  # extract the completed matrix
  UImatrix_filled <- as.matrix(svt_result$X)
  # calculate the RMSE between the original matrix and the completed matrix
  rmse <- sqrt(mean((UImatrix_sub - UImatrix_filled)^2, na.rm = TRUE))
  # add the results to the data frame
  results <- results %>% add_row(lambda = lambda, rmse = rmse)
}

# find the lambda value with the lowest RMSE
best_lambda <- results %>% filter(rmse == min(rmse)) %>% pull(lambda)

best_lambda
```

```{r, cache=TRUE}
# full matrix
# SVT <- fill.SVT(UImatrix, lambda = 1, maxiter = 50, tol = 0.01)
```

```{r, cache=TRUE}
# subset matrix using best_lambda
SVTsubBest <- fill.SVT(UImatrix_sub, lambda = 0.001, maxiter = 50, tol = 0.001)

head(sort(SVTsubBest$X, decreasing = TRUE))
SVTsub_matrix <- as.matrix(SVTsubBest$X)
```


# Recommonder system 
## Cosine simlarity

```{r}
library(lsa)

# compute cosine similarity between users
cos_sim <- cosine(as.matrix(SVTsub_matrix))

colnames(cos_sim) <- movie_titles

diag(cos_sim) <- 0
```

```{r}
# function to get top n similar items for a given item
get_top_n_items <- function(user, num) {
  n <- num
  # get the similarity scores for the item of interest
  scores <- sort(cos_sim[user, ], decreasing = TRUE)
  # get the indices of the top n similar items
  top_n_indices <- names(scores)[1:n]
  # get the corresponding scores
  top_n_scores <- scores[1:n]
  return(data.frame(item = top_n_indices, score = top_n_scores))
}

# example usage: get top 13 recommended items for user 3
top_n_items <- get_top_n_items(3, 13)
top_n_items
```














